---
title: "Green sea turtle analysis "
author: "Ketong Zhang"
output: html_document
date: "2024-03-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(naniar)
```

**load the dataset and clean names**
```{r}
green_turtle <- read_csv("data/Green_sea_turtle_age_growth.csv") %>% clean_names()
```

**take a look at the dataset**
```{r}
glimpse(green_turtle)
```

**check if there are any NAs**
```{r}
miss_var_summary(green_turtle)
```

**produce a summary of average straightline carapace length of turtle**
```{r}
green_turtle %>% 
  group_by(sex) %>% 
  filter(straightline_carapace_length != "NA") %>% 
  summarize((avg_length = mean(straightline_carapace_length)))
```

**produce a graph on the difference in straightline carapace length between male and female**

female turtles are larger than male turtles 
```{r}
green_turtle %>%
  group_by(sex) %>%
  filter(straightline_carapace_length != "NA") %>% 
  summarize(avg_length = mean(straightline_carapace_length)) %>%
  ggplot(aes(x = sex, y = avg_length, fill = sex)) +
  geom_col(color = "black", fill = "lightskyblue3") +
  theme_light()+
  labs(title = "Average Straightline Carapace Length by Sex",
       x = "Sex",
       y = "Average Length",
       fill = "Sex")+
  theme(plot.title=element_text(size=rel(1.75), hjust = 0.5))
```

**compare growth rate by year**
based on the result, you can see there is an overall negative correlation
there is a growth between age 5 to 12, and follow by decrease in growth rate after the age of 12
```{r}
green_turtle %>% 
  group_by(age) %>% 
  summarize(avg_growth = mean(growth_rate, na.rm = TRUE)) %>% 
  ggplot(aes(x = age, y = avg_growth, color = age))+
  geom_line()+
  labs(title = "Growth Rate by Age",
       x = "Age of turtle",
       y = "Average growth rate",
       fill = "Age")+
  theme(plot.title=element_text(size=rel(1.75), hjust = 0.5))
```

**the most "popular" age of stranded turtle**
female = 8
male = 7
reason: around 7-8, the juvenile stage of turtles are likely to be more active, therefore increase the stranding possibilities
```{r}
green_turtle %>% 
  group_by(sex) %>% 
  count(age) %>% 
  arrange(desc(n)) %>% 
  ggplot(aes(x = age, y = n))+
  geom_col(position = "dodge", alpha = 0.9, fill = "olivedrab3", color = "salmon4")+
  geom_text(aes(label = age), vjust = -0.5, color = "black", size = 2)+
  facet_wrap(~sex)+
  labs(title = "Turtle stranding ages",
       x = "Turtle's age",
       y = "Number of stranding",
       fill = "Sex")+
  theme(plot.title=element_text(size=rel(1.75), hjust = 0.5))+
  scale_x_discrete()
```

**the most popular year of stranding**
(top 5 year of stranding)
scale the year to discrete
```{r}
green_turtle %>% 
  count(year) %>% 
  arrange(desc(n)) %>% 
  ggplot(aes(x = as.factor(year), y = n))+
  geom_col(fill = "palegreen3")+
  labs(title = "Green Sea Turtle Stranding Trend Over The Years",
       x = "Years",
       y = "Number of stranding")+
  theme(plot.title=element_text(size=rel(1.75), hjust = 0.5))+
  scale_x_discrete(limits=c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009"))
```

**range / boxplot on the range of length by sex**
```{r}
green_turtle %>% 
  group_by(sex) %>% 
  ggplot(aes(x = sex, y = straightline_carapace_length))+
  geom_boxplot(fill = "lightblue3")+
  labs(title = "Range of Straightline Carapace Length",
       x = "Sex",
       y = "Straightline carapace length")+
  theme(plot.title=element_text(size=rel(1.75), hjust = 0.5))
```

```{r}
green_turtle %>% 
  filter(sex == "Male") %>% 
  summary(straightline_carapace_length)
```

```{r}
green_turtle %>% 
  filter(sex == "Female") %>% 
  summary(straightline_carapace_length)
```

