---
title: "adult loggerhead turtles"
author: "Amanda Duong"
date: "2024-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
library("naniar")
library("ggplot2")
library("dplyr")
```

# clean the data first
```{r}
loggerheads <- read_csv("data/Adult_loggerhead_turtle.csv") %>% clean_names()
```
# look at column names
```{r}
names(loggerheads)
```
# look at summary of data
```{r}
glimpse(loggerheads)
```

# Graph of Average Growth over Time Per State
```{r}
loggerheads %>%
  count(state) # first get list of states in data
```

```{r}
loggerheads <- loggerheads %>% # combine FL values into one
  mutate(state = case_when(
    state %in% c("FL", "FL(FA)", "FL (FA)") ~ "FL",
    TRUE ~ state
  ))
```

```{r}
loggerheads_long <- loggerheads %>%
  select(state, year, growth) %>%
  group_by(state, year) %>%
  summarize(mean_growth = mean(growth, na.rm = T))
```

```{r}
# Creating a line plot of average growth over time for all states
loggerheads_long %>%
  ggplot(aes(x = year, y = mean_growth, color = state)) +
  geom_line() +
  labs(title = "Mean Growth for Each State Over the Years",
       x = "Year", y = "Mean Growth Rate") +
  theme_minimal()
```
# Do males or females tend to grow faster over time? (Removing unknowns)
```{r}
loggerheads %>%
  select(sex, year, growth) %>%
  filter(sex!="NA") %>%
  filter(sex!="unknown") %>%
  group_by(sex, year) %>%
  summarize(mean_growth = mean(growth, na.rm = T)) %>%
  ggplot(aes(x = year, y = mean_growth, color = sex)) +
  geom_line() +
  labs(title = "Mean Growth for Each Sex Over the Years",
       x = "Year", y = "Mean Growth Rate") +
  theme_minimal()
```

# Is there a correlation between age and growth rate?
```{r}
loggerheads %>%
  select(sex, growth) %>%
  filter(sex!="unknown") %>%
  group_by(sex) %>%
  summarise(mean_growth = mean(growth, na.rm=T)) %>%
  arrange(desc(mean_growth))
```
# add color to boxplot
```{r}
loggerheads %>%
  filter(sex != "unknown") %>%
  ggplot(aes(x = sex, y = growth)) +
  geom_boxplot() +
  scale_y_log10() +  # Adding logarithmic scale
  labs(title = "Boxplot of Logarithmic Growth by Sex",
       x = "Sex",
       y = "Logarithmic Growth") +
  theme_minimal()
```
# Females tend to grow in size faster than males. 

# At what age do males or females tend to grow the fastest?
```{r}
loggerheads %>%
  select(sex, age, growth) %>%
  filter(sex!="NA") %>%
  filter(sex!="unknown") %>%
  group_by(sex, age) %>%
  summarize(mean_growth = mean(growth, na.rm = T)) %>%
  ggplot(aes(x = age, y = mean_growth, color = sex)) +
  geom_line() +
  labs(title = "Mean Growth Rate for Each Sex with Age",
       x = "Age", y = "Mean Growth") +
  theme_minimal()
```

# show popular ages of males and females in data
```{r}
loggerheads %>% 
  group_by(sex) %>% 
  filter(sex != "unknown") %>%
  count(age) %>% 
  arrange(desc(n)) %>% 
  ggplot(aes(x = age, y = n)) +
  geom_col(position = "dodge", alpha = 0.5, fill = "pink", color = "plum") +
  geom_text(data = . %>% slice_max(n, with_ties = FALSE), aes(label = age),
            position = position_dodge(width = 0.9), vjust = -0.5, color = "black", size = 2) +
  facet_wrap(~sex) +
  labs(title = "Turtle Ages by Sex",
       x = "Age",
       y = "Count",
       fill = "Sex") +
  theme(plot.title = element_text(size = rel(1.75), hjust = 0.5))
```
# count number of females and males in data, removing unknowns
```{r}
loggerheads %>% 
  group_by(sex) %>% 
  filter(sex != "unknown") %>%
  count(sex) 
```

