---
title: "Sea Turtle Strandings Florida"
author: "Jasmine Hsiao"
date: "2024-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
library("naniar")
library(ggplot2)
library(ggmap)
library(RColorBrewer)
```

## Load the data
```{r message=FALSE, warning=FALSE}
florida_stranding <- read_csv("data/Sea_Turtle_Strandings_Florida.csv") %>% clean_names()
```

## Structure and NAs
```{r}
summary(florida_stranding)
```

```{r}
glimpse(florida_stranding)
```

```{r}
florida_stranding %>% 
  naniar::miss_var_summary()
```
### Change x and y to long and lat
```{r}
florida_stranding <- florida_stranding %>% 
  rename(longitude = x, latitude = y)
```

## Where are the sea turtle strandings by species?
### Build a map
```{r}
register_stadiamaps("cf7e4fa6-eb1a-4afb-a4a8-6d5d47e2a05a", write = FALSE) #API Key
```

```{r}
florida_stranding %>% 
  #y=latitude, x=longitude
  select(longitude, latitude) %>% 
  summary()
```

```{r}
lat_florida <- c(24.23, 30.72)
long_florida <- c(-87.52, -79.49)
bbox_florida <- make_bbox(long_florida, lat_florida, f = 0.03) #f is the fraction of the bounding box to add to the range
```

```{r}
map_florida <- get_stadiamap(bbox_florida, maptype = "stamen_terrain", zoom=7)
```

```{r}
green_palette <- RColorBrewer::brewer.pal(7, "PRGn")

ggmap(map_florida) + 
  geom_point(data = florida_stranding, aes(longitude, latitude, color = commonname), 
             size = 0.5) +
  scale_color_manual(values = setNames(green_palette, levels(florida_stranding$commonname))) +
  labs(x = "Longitude", y = "Latitude", title = "Sea Turtle Stranding Locations in Florida by Species")
```
## What are the total counts of sea turtle strandings in the data?
```{r}
florida_stranding %>% 
  group_by(commonname) %>% 
  count(commonname) %>% 
  arrange(desc(n))
```
```{r}
florida_stranding %>% 
  count(commonname) %>%
  ggplot(aes(x = reorder(commonname, -n), y = n, fill = commonname)) +
  geom_col(alpha = 0.7) +
  geom_text(aes(label = n), vjust = -0.5, color = "grey43", size = 3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_text(size = rel(1), hjust = 0.5)) +
  labs(title = "Total Number of Sea Turtle Strandings by Species in Florida Through the Years",
       y = "Number of Strandings",
       x = "Species") +
  scale_fill_manual(values = rep("darkgreen", n_distinct(florida_stranding$commonname)))+
  guides(fill = FALSE)
```
## How has the number of sea turtle in Florida changed throughout the years?
```{r}
florida_stranding %>% 
  mutate(year = as.factor(year),
         alternating_color = ifelse(as.numeric(year) %% 2 == 0, "#006E2E", "#57BB8A")) %>% 
  ggplot(aes(x = year, fill = alternating_color)) +
  geom_bar(alpha = 0.7) +
  scale_fill_manual(values = c("#006E2E", "#57BB8A")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(size = rel(1), hjust = 0.5)) +
  labs(title = "Number of Sea Turtle Strandings by Year in Florida (1980-2022)",
       y = "Number of Strandings") +
  guides(fill = FALSE)
```


## Is there a specific month where different species of sea turtles are more likely to get stranded?
```{r}
florida_stranding %>% 
  ggplot(aes(x=month))+
  geom_bar(fill="#C3B1E1", alpha=0.9)+
  facet_wrap(~commonname)+ 
  theme(axis.text.x = element_text(angle = 60, hjust=1),
        plot.title=element_text(size=rel(1), hjust = 0.5))+
  labs(title="Number of Sea Turtle Strandings by Species in Each Month in Florida",
       y="Number of Strandings",
       x="Month")+
  scale_x_discrete(limits=c("1","2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
```

## How does the condition of sea turtle stranding compare between green turtle and loggerhead? 
```{r}
green_palette <- c("#D3FCC3", "#57BB8A")

florida_stranding %>% 
  filter(commonname %in% c("Green turtle", "Loggerhead")) %>% 
  filter(condition != "NULL") %>% 
  group_by(commonname, condition) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = condition, y = count, fill = commonname)) + 
  geom_col(position = "dodge", alpha = 0.8) +
  geom_text(aes(label = count), position = position_dodge(width = 0.8), vjust = -0.5, color = "grey43", size = 2.5) +
  scale_fill_manual(values = setNames(green_palette, c("Green turtle", "Loggerhead"))) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_text(size = rel(1), hjust = 0.5)) +
  labs(title = "Strandings Condition of Green Turtle and Loggerhead in Florida",
       y = "Number of Strandings",
       x = "Condition",
       fill = "Species")
```



